The Weakly Nonlinear Magnetorotational Instability

### 2016-08-31 paper/widegap_mri_paper.tex
store growth rate and frequency in linear eigenvalue and adjoint objects, and in run output

A real GLE has also been found to describe the formation of zonal flows out of magnetized turbulence in a model system (\citei{Parker:2013hy}). Zonal flows are axisymmetric structures, large-scale and long-lived, which form spontaneously out of turbulence. They have recently been observed in some numerical studies of the MRI, and have generated considerable interest for their possible role in planet formation in protoplanetary disks (\citei{Johansen:2009uj}, \citei{Kunz:2013}). The present work is of course an idealized geometry, and we make no attempt to model a realistic protoplanetary disk environment. However, it is worth noting that the GLE we derive implies that axisymmetric, large-scale, long-lived structures are a generic feature of the MRI in the weakly nonlinear regime. This work provides a mathematical description of the MRI as a pattern-forming process, but much remains to be understood, particularly involving the relevance of these insights to realistic astrophysical disks. Our model is most directly relevant to Taylor-Couette flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.

>>>>

A real GLE has also been found to describe the formation of zonal flows out of magnetized turbulence in a model system (\citei{Parker:2013hy}). Zonal flows are axisymmetric structures, large-scale and long-lived, which form spontaneously out of turbulence. They have recently been observed in some numerical studies of the MRI, and have generated considerable interest for their possible role in planet formation in protoplanetary disks (\citei{Johansen:2009uj}, \citei{Kunz:2013}). The present work is of course an idealized geometry, and we make no attempt to model a realistic protoplanetary disk environment. However, it is worth noting that the GLE we derive implies that axisymmetric, large-scale, long-lived structures are a generic feature of the MRI in the weakly nonlinear regime. This work provides a mathematical description of the MRI as a pattern-forming process, but much remains to be understood, particularly involving the application of this model system to realistic astrophysical disks. Our model is most directly relevant to Taylor-Couette flows, and we emphasize that laboratory MRI experiments stand poised to observe the MRI-driven pattern formation predicted here.
